\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities

### 2020-12-27 popl-2019-journal-v/paper.tex
Full pass over paper, fix another reviewer remark.

In this work, we propose and study \stktokens{}: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997,szabo_scarce_objects} and in technical documents.
Concurrently with our work, Watson et al. have developed a (more realistic) design for linear capabilities in CHERI that is detailed in the latest CHERI ISA reference~\citep{watson_capability_2020}.
% KJAA: Perhaps make it clear by who and when/where?
The hardware prevents these capabilities from being duplicated.
We propose to make stack and return pointers linear and make components hand them out in cross-component invocations and require them back on return.
The non-duplicability of linear capabilities together with some security checks allows us to guarantee WBCF and LSE without large overhead on boundary crossings and in particular without the need for clearing large blocks of memory.
To avoid confusion, it is worth pointing out that our linear caabilities are really affine rather than linear: erasing them is allowed.
However, we have chosen to use the term ``linear'' to align with common usage in the world of linear type systems.

@@

\section{A Capability Machine with Sealing and Linear Capabilities}
\label{sec:cap-mach-w-seal-and-lin}
% Brief intro to section and this cap machine:
% Cap machine inspired by CHERI, but with linear capabilities
In this section, we introduce a simple but representative capability machine with linear capabilities, that we call \trgcm{} (Linear Capability Machine).
\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri_2015} with linear capabilities as the main addition.
For simplicity, \trgcm{} assumes an infinite address space and unbounded integers.
% \lau{Do we have a good reason to assume an infinite address space for this work?
%   We do not model malloc, so we do not need to worry about what happens when malloc runs out of memory.}

%%% Capabilities
The concept of a capability is the cornerstone of any capability machine.
In its simplest form, a capability is a permission and a range of authority.
The permission dictates the operations the capability can be used for, and the range of authority specifies the range of memory it can act upon.
% Permissions and Range of authority
The capabilities on \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (defined in Figure~\ref{fig:target-syntax} with the rest of the syntax of \trgcm{}).
Here $\var{perm}$ is the permission, and $[\var{base},\var{end}]$ is the range of authority.
The available permissions are read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$) ordered by $\le$ as illustrated in Figure~\ref{fig:perm-hier}.
% Pointer and linearity
In addition to the permission and range, capabilities also have a current address $\var{addr}$ and a linearity $\var{lin}$.
The current address is a CHERI design choice that makes it easier to implement pointers as capabilities in C-like languages \cite{woodruff_cheri_2014}.
The linearity is either $\normal$ for traditional capabilities or $\linear$ for linear ones.
% What is linear capabilities, split/splice
A linear capability is a capability that cannot be duplicated.
This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or memory, the source is cleared.
The non-duplicability of linear capabilities means that a linear capability cannot become aliased if it wasn't to begin with.

@@

The permission to make sealed capabilities is represented by a second form of capability in addition to the memory capabilities we saw above: a set-of-seals capability $\seal{\sigma_\var{base},\sigma_\var{end},\sigma_{\var{current}}}$.

@@

% load w/ lin (and store)
The \texttt{store} and \texttt{load} instructions are fairly standard.
They require a capability with permission to either write or read depending on the operation, they check that the capability points within the range of authority.
Linear capabilities introduce one extra complication for \texttt{load} as it needs to clear the loaded memory address when it contains a linear capability in order to not duplicate the capability.
In this case, we require that the memory capability used for loading also has write-permission.

@@

%% Sealing
Two instructions manipulate seals in \trgcm{}: \texttt{cseal} for sealing a capability and \texttt{xjmp} for unsealing a pair of capabilities.
% cseal
Given a sealable $\vsc$ and a set-of-seals where the current seal $\sigma$ is within the range of available seals, the \texttt{cseal} instruction seals $\vsc$ with $\sigma$.
% xjmp
Apart from dealing with linearity, \texttt{xjmp} takes a pair of sealed capabilities, unseals them, and puts one in the $\pcreg$ register and the other in the $\rdata$ register, but only if they are sealed with the same seal and the data capability (the one placed in $\rdata$) is non-executable.
A pair of sealed capabilities can be seen as a closure where the code capability (the capability placed in $\pcreg$) is the program and the data capability is the local environment.
Because of the opacity of sealed capability, the creator of the closure can be sure that execution will start where the code capability points and only in an environment with the related data, i.e.\ sealed with the same seal.
This makes \texttt{xjmp} the mechanism on \trgcm{} that transfers control between security domains.
Opaque sealed capabilities encapsulate a security domain's local state and authority, and they only become accessible again when control is transferred to the security domain.
Some care should be taken for sealing because reusing the same seal for multiple closures makes it possible to jump to the code of one closure with the environment of another.
% xjmp to unseal
\trgcm{} does not have an instruction for unsealing capabilities directly, but we will explain below how it can be (partially) simulated using \texttt{xjmp}.

% split/splice
Instructions for reducing the authority of capabilities are commonplace on capability machines as they allow us to limit what a capability can do before it is passed away.
For normal capabilities, reduction of authority can be done without actually giving up any authority by duplicating the capability first.
With linear capabilities authority cannot be preserved in this fashion as they are non-duplicable.
In order to make a lossless reduction of the range of authority, \trgcm{} provides special hardware support in the form of \texttt{split} and \texttt{splice}.
The \texttt{split} instruction takes a capability with range of authority $[\var{base},\var{end}]$ and an address $n$ and creates two new capabilities, with $[\var{base},n]$ and $[n+1,\var{end}]$ as ranges of authority.
Everything else, i.e.\ permission, linearity and current address, is copied without change to the new capabilities.
With \texttt{split}, we can reduce the range of authority of a linear capability without losing any authority as we retain it in the second capability.
The \texttt{splice} instruction essentially does the inverse of \texttt{split}.
Given two capabilities with adjacent ranges of authority and the same permissions and linearity, \texttt{splice} splices them together into one capability.
The two instructions work in the same way for seal sets.
We do not provide special support for lossless reduction of capability permissions, but this could probably be achieved with more fine-grained permissions.
This would also allow linear capabilities to have aliases, but only by linear capabilities with disjoint permissions.

\begin{jversion}
% The remaining instructions
  The interpretation of the remainder of the instructions are displayed in Appendix~\ref{app:instr-interpretation}.
  The instructions $\tgetb{}{}$, $\tgete{}{}$, $\tgetlin{}{}$, and $\tgetp{}{}$ all project information about capabilities.
  The $\tgetb{}{}$ and $\tgete{}{}$ instructions, respectively, project the base and end address of the range of authority.
  The linearity of a permission is projected with $\tgetlin{}{}$, and finally the permission is projected with $\tgetp{}{}$.
  The instructions $\tgetb{}{}$ and $\tgete{}{}$ also work on sets of seals.
  The instruction $\tisptr{}{}$ returns an integer representation of the type of a word (capability or number).
  \trgcm{} also has arithmetic instructions $\tplus{}{}{}$, $\tminus{}{}{}$, and $\tlt{}{}{}$.
  The latter instruction compares two numbers and writes 1 or 0 to a target register depending on whether one number is less than the other.
  The instruction $\tsetatob{}$ sets the current address (or seal) of a capability (or set-of-seals) to the base address of the range of authority (or range of seals).
  This instruction makes it easy to work relatively to the base address of a capability.
  This instruction is not strictly necessary as it can be emulated with other instructions.
  Finally, we have the $\trestrict{}{}$ instruction which restricts the permission of a capability according to the $\le$ relation.
\end{jversion}

@@

  Additionally, when the untrusted code returns,  we need to regain access to the capabilities that represent our authority, which we had access to before passing control.
  To achieve this, we have to store them in a location that we can access after the return.
  However, without an encapsulation mechanism, the untrusted code would have access to all this through the return capability because it gives the same authority to the untrusted code as to us after the return.
  This is why any reasonable capability machine must have an encapsulation mechanism to allow programs to set up boundaries between security domains.

  In \trgcm{}, we can seal the return capability to establish such a boundary.
  Specifically, we can seal the return capability (which points to the instruction to be executed after the return) as a pair, together with a pointer to the stack frame where we have stashed away our capabilities (see further).
  By doing this, the untrusted code is prevented from changing the target of the return capability forcing them to return to the point we specified.
  Additionally, they do not get access to the capabilities we stashed away.
  Nevertheless, the automatic unsealing of sealed capability pairs upon invocation will ensure that the return code can still proceed as before.
  In other words, the sealing mechanism in \trgcm{} allows us to transfer control to untrusted code without giving up our capabilities or handing them over, and also control the locations in our code they can jump back to.

  It is worth pointing out that \trgcm{} does not have a direct unsealing instruction to extract a capability from its sealed version when the seal is available, but one can be emulated.
  Say you have a sealed non-executable word $\sealed{\sigma,w}$ as well as a set-of-seals capability $\seal{\sigma_\baddr,\sigma_\eaddr, \sigma}$ that contains $\sigma$, i.e.\ $\sigma \in [\sigma_\baddr,\sigma_\eaddr]$.
  It is possible to extract $w$ in $\rdata$ using $\txjmp{}{}$ in the following way.
  The idea is to use the seal to construct a sealed version of the pc capability incremented by one, and then perform an $\txjmp{}{}$ to it in combination with the sealed capability.
  This does not work for sealed executable capabilities because $\txjmp{}{}$ fails if the data capability is executable.
  However, even though we cannot extract the executable capability from $\sealed{\sigma,c}$, we can still invoke it with arbitrary arguments, since we can use the seal to construct an arbitrary data part and $\txjmp{}{}$ to the combination.\footnote{Alternatively, we could remove the current restriction in \texttt{xjmp} that requires the data capability to be non-executable but then teh invoked code would often have to manually perform this check instead.}

@@

\begin{jversion}
\subsection{Decoding and encoding functions}
The operational semantics of the capability machine uses the function $\decInstr{}$ to decode instructions.
We also assume a function $\encInstr{}$ to make it easy to specify programs in terms of instructions.
Rather than defining such a decode function and an encode function, we assume that they are given with certain properties.
The $\decInstr{} : \Word \fun \Instr$ should decode all capabilities to the $\fail$ instruction, i.e.\ only numbers are decoded to non-fail instructions:
\[
\forall c \in \Caps \ldotp \decInstr{c} = \fail
\]
The $\decInstr{}$ should have a right inverse $\encInstr{} : \Instr \fun \ints$, i.e.
\[
\forall i \in \Instr \ldotp \decInstr{}(\encInstr{i}) = i
\]
These assumptions are sufficient to construct program examples in terms of instructions rather than machine words (using $\encInstr{}$), and run them on the machine (using the fact that $\decInstr{}$ is the left inverse of $\encInstr{}$).
% The $\encInstr{}$ function allows us to specify the program ``abstractly''.
% As the $\decInstr{}$ function is the left inverse of $\encInstr{}$, we can even execute the program in the operational semantics without ever worrying about what the actual encoding is.
%This is not to say that encoding is an easy problem. There are many things to get.

@@

\subsection{Components, linking, programs, and contexts}
\label{subsec:components-linking}
The executable configuration describes the machine state, but it does not make it clear what components run on the machine and how they interact with each other.
To clarify this, we introduce notions of components and programs from which we construct executable configurations.
A component (defined in Figure~\ref{fig:target-component-and-linking}) is basically a program with entry points in the form of imports that need to be linked.
It has exports that can satisfy the imports of other components.
% Structure
A base component $\comp_0$ consists of a code memory fragment, a data memory fragment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals (see Section~\ref{sec:form-secur-with}), and a set of all the linear addresses (addresses governed by a linear capability).
The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.
An export associates a word with a symbol.
%base = library
A component is either a library component (without a main entry point) or an incomplete program with a main in the form of a pair of sealed capabilities.
The latter can be seen as a program that still needs to be linked with libraries.
% Linking .
Components are combined into new components by linking them together, as long as only one has a main function.
Two components can be linked when their memories, seals, and linear addresses are disjoint.
They are combined by taking the union of each of their constituents.
For every import that is satisfied by an export of the other component, the data memory is updated to have the exported word on the imported address.
The satisfied imports are removed from the import list in the resulting linked component and the exports are obtained by combining the components' exports.

% Programs
We can now define the notion of a program as well as a context.
\begin{definition}[Programs and Contexts]
  \label{def:program-and-context}
  A \emph{program} is a component $(\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with an empty import list.
  A \emph{context} for a component $\comp$ is a component $\comp'$ such that $\comp \bowtie \comp'$ is a program.
\end{definition}
% Getting an executable from a component - postpone this to later, but mention:
% write xor exec
How a program is initialised to create an executable configuration is discussed in Section~\ref{sec:form-secur-with}.
% We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
%Further, we will only consider a specific class of well-formed components, but we will get back to this.
Some simplifications have been made in this presentation of \trgcm{}. See \citet{technical_report_popl} for details.
\begin{figure}[tb]
  \centering
  \[
    \begin{array}{rl c rl c rl}
    s &\in \Symbol & \phantom{make} &
    \mathrm{import} &\mathrel{::=} a \mapsfrom s & \phantom{make} &
    \mathrm{export} &\mathrel{::=} s \mapsto w\\
    \comp_0 & \multicolumn{7}{l}{\mathrel{::=} (\mscode,\msdata,\overline{\mathrm{import}},\overline{\mathrm{export}},\sigrets,\sigcloss,A_\linear)}\\
    \comp & \multicolumn{7}{l}{\mathrel{::=} \comp_0 \mid  (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})}
  \end{array}
\]
\vspace{0.5cm}
\begin{mathpar}
  \inference{
    \comp_0 = (\mscode[1], \msdata[1], \overline{\var{import}_1}, \overline{\var{export}_1}, \sigrets[1], \sigcloss[1],A_{\linear,1})\\
    \comp_0' = (\mscode[2], \msdata[2], \overline{\var{import}_2}, \overline{\var{export}_2}, \sigrets[2], \sigcloss[2],A_{\linear,2})\\
    \comp_0'' = (\mscode[3], \msdata[3], \overline{\var{import}_3}, \overline{\var{export}_3}, \sigrets[3], \sigcloss[3],A_{\linear,3})\\
    \mscode[3] = \mscode[1] \uplus \mscode[2] \\
    \msdata[3] = (\msdata[1] \uplus \msdata[2])[a \mapsto w \mid (a \mapsfrom s) \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}), (s \mapsto w) \in \overline{\var{export}}_3] \\
    \overline{\var{export}_3} = \overline{\var{export}_1} \cup \overline{\var{export}_2}&
    \overline{\var{import}_3} = \{ a \mapsfrom s \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}) \mid s \mapsto \_ \not\in \overline{\var{export}_3} \}\\
    \sigrets[3] = \sigrets[1] \uplus \sigrets[2] &
    \sigcloss[3] = \sigcloss[1] \uplus \sigcloss[2] &
    A_{\linear,3} = A_{\linear,1} \uplus A_{\linear,2}\\
    \dom(\mscode[3]) \mathrel{\#} \dom(\msdata[3]) & \sigrets[3] \mathrel{\#} \sigcloss[3]
  } {
    \comp_0'' = \comp_0 \bowtie \comp_0'
  }
  \and
  \inference{
    \comp_0'' = \comp_0 \bowtie \comp_0'
  }{
    \begin{aligned}
      (\comp_0'',c_{\mathrm{main},c}, c_{\mathrm{main},d}) &= \comp_0 \bowtie
      (\comp_0',c_{\mathrm{main},c}, c_{\mathrm{main},d})\\
      &=
      (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}) \bowtie \comp_0'
    \end{aligned}
  }
  \end{mathpar}
% \begin{definition}[Plugging a program into a context]
%   When $\var{comp'}$ is a context for component $\comp$ and $\comp' \bowtie \comp \rightsquigarrow \Phi$, 
%   then we write $\plug{\comp'}{\comp}$ for the execution configuration $\Phi$.
% \end{definition}
  \caption{Components and linking of components.}
  \label{fig:target-component-and-linking}
\end{figure}

@@

% Informally explain how we prevent this using linear capabilities
As the examples illustrate, this naive use of memory and object capabilities does not suffice to enforce LSE and WBCF.
The problem is essentially that the stack and return pointers that a callee receives from a caller remain in effect outside their intended lifetime: either when the callee has already returned or when they have themselves invoked other code. 
Linear capabilities offer a form of revocation\footnote{Revocation in the sense that if we hand out a linear capability and later get it back, then the receiver cannot have kept a copy of it as it is non-duplicable.} that can be used to prevent this from happening.

>>>>

In this work, we propose and study \stktokens{}: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997,szabo_scarce_objects} and in technical documents.
Concurrently with our work, \citeauthor{watson_capability_2020} have developed a (more realistic) design for linear capabilities in CHERI that is detailed in the latest CHERI ISA reference~\citep{watson_capability_2020}.
% KJAA: Perhaps make it clear by who and when/where?
The hardware prevents these capabilities from being duplicated.
We propose to make stack and return pointers linear and make components hand them out in cross-component invocations and require them back on return.
The non-duplicability of linear capabilities together with some security checks allows us to guarantee WBCF and LSE without large overhead on boundary crossings and in particular without the need for clearing large blocks of memory.
To avoid confusion, it is worth pointing out that our linear capabilities are really affine rather than linear: erasing them is allowed.
However, we have chosen to use the term ``linear'' to align with common usage in the world of linear type systems.

@@

\section{A Capability Machine with Sealing and Linear Capabilities}
\label{sec:cap-mach-w-seal-and-lin}
% Brief intro to section and this cap machine:
% Cap machine inspired by CHERI, but with linear capabilities
In this section, we introduce a simple but representative capability machine with linear capabilities, that we call \trgcm{} (Linear Capability Machine).
\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri_2015} with linear capabilities as the main addition.
For simplicity, \trgcm{} assumes an infinite address space and unbounded integers.\footnote{Although we have not thoroughly investigated, we do not believe our results depend heavily on these assumptions and we expect they could be lifted without a great impact on the proofs.}
% \lau{Do we have a good reason to assume an infinite address space for this work?
%   We do not model malloc, so we do not need to worry about what happens when malloc runs out of memory.}

%%% Capabilities
The concept of a capability is the cornerstone of any capability machine.
In its simplest form, a (memory) capability is a permission and a range of authority.
The former dictates the operations the capability can be used for, and the latter specifies the range of memory it can act upon.
% Permissions and Range of authority
Memory capabilities on \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (defined in Figure~\ref{fig:target-syntax} with the rest of the syntax of \trgcm{}).
Here $\var{perm}$ is the permission, and $[\var{base},\var{end}]$ is the range of authority.
The available permissions are read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$) ordered by $\le$ as illustrated in Figure~\ref{fig:perm-hier}.
% Pointer and linearity
In addition to the permission and range, capabilities also have a current address $\var{addr}$ and a linearity $\var{lin}$.
The current address is a CHERI design choice that makes it easier to implement pointers as capabilities in C-like languages \cite{woodruff_cheri_2014}.
The linearity is either $\normal$ for traditional capabilities or $\linear$ for linear ones.
% What is linear capabilities, split/splice
A linear capability is a capability that cannot be duplicated.
This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or memory, the source is cleared.
The non-duplicability of linear capabilities means that a linear capability cannot become aliased if it wasn't to begin with.

@@

The permission to make sealed capabilities is represented by a second form of capability (in addition to the memory capabilities we saw above): a set-of-seals capability $\seal{\sigma_\var{base},\sigma_\var{end},\sigma_{\var{current}}}$.

@@

% load w/ lin (and store)
The \texttt{store} and \texttt{load} instructions are fairly standard.
They require a capability with permission to write or read (respectively), they check that the capability points within the range of authority.
Linear capabilities introduce one extra complication for \texttt{load} as it needs to clear the loaded memory address when it contains a linear capability in order to not duplicate the capability.
In this case, we require that the memory capability used for loading also has write-permission.

@@

%% Sealing
Two instructions manipulate seals in \trgcm{}: \texttt{cseal} for sealing a capability and \texttt{xjmp} for unsealing a pair of capabilities.
% cseal
Given a sealable $\vsc$ and a set-of-seals where the current seal $\sigma$ is within the range of available seals, the \texttt{cseal} instruction seals $\vsc$ with $\sigma$.
% xjmp
Apart from dealing with linearity, \texttt{xjmp} takes a pair of sealed capabilities, unseals them, and puts one in the $\pcreg$ register and the other in the $\rdata$ register, but only if they are sealed with the same seal and the data capability (the one placed in $\rdata$) is non-executable.
A pair of sealed capabilities can be seen as a closure where the code capability (the capability placed in $\pcreg$) is the program and the data capability is the local environment.
Because of the opacity of sealed capabilities, the creator of the closure can be sure that execution will start where the code capability points and only in an environment with the related data, i.e.\ with access to a data capability sealed with the same seal.
This makes \texttt{xjmp} the mechanism on \trgcm{} that transfers control between security domains.
Opaque sealed capabilities encapsulate a security domain's local state and authority, and they only become accessible again when control is transferred to the security domain.
Some care should be taken for sealing because reusing the same seal for multiple closures makes it possible to jump to the code of one closure with the environment of another.
% xjmp to unseal
\trgcm{} does not have an instruction for unsealing capabilities directly, but we will explain below how it can be (partially) simulated using \texttt{xjmp}.

% split/splice
Instructions for reducing the authority of capabilities are commonplace on capability machines as they allow us to limit a capability's authority before passing it to other code.
For normal capabilities, reduction of authority can be done without actually giving up any authority by duplicating the capability first.
With linear capabilities, authority cannot be preserved in this fashion as they are non-duplicable.
In order to make a lossless reduction of the range of authority, \trgcm{} provides special hardware support in the form of \texttt{split} and \texttt{splice}.
The \texttt{split} instruction takes a capability with range of authority $[\var{base},\var{end}]$ and an address $n$ and creates two new capabilities, with $[\var{base},n]$ and $[n+1,\var{end}]$ as ranges of authority.
Everything else, i.e.\ permission, linearity and current address, is copied without change to the new capabilities.
With \texttt{split}, we can reduce the range of authority of a linear capability without losing any authority as we retain it in the second capability.
The \texttt{splice} instruction essentially does the inverse of \texttt{split}.
Given two capabilities with adjacent ranges of authority and the same permissions and linearity, \texttt{splice} splices them together into one capability.
The two instructions work in the same way for set-of-seals capabilities.
We do not provide special support for lossless reduction of capability permissions, but this could probably be achieved with more fine-grained permissions.
This would also allow linear capabilities to have aliases, but only by linear capabilities with disjoint permissions.

\begin{jversion}
% The remaining instructions
  The interpretation of the remainder of the instructions are displayed in Appendix~\ref{app:instr-interpretation}.
  The instructions $\tgetb{}{}$, $\tgete{}{}$, $\tgetlin{}{}$, and $\tgetp{}{}$ all query information about capabilities.
  The $\tgetb{}{}$ and $\tgete{}{}$ instructions, respectively, project the base and end address of the range of authority.
  The linearity of a permission is projected with $\tgetlin{}{}$, and finally the permission is projected with $\tgetp{}{}$.
  The instructions $\tgetb{}{}$ and $\tgete{}{}$ also work on set-of-seal capabilities.
  The instruction $\tisptr{}{}$ returns an integer representation of the type of a word (capability or number).
  \trgcm{} also has arithmetic instructions $\tplus{}{}{}$, $\tminus{}{}{}$, and $\tlt{}{}{}$.
  The latter instruction compares two numbers and writes 1 or 0 to a target register depending on whether one number is less than the other.
  The instruction $\tsetatob{}$ sets the current address (or seal) of a memory (or set-of-seals) capability to the base address of the range of authority (or range of seals).
  This instruction makes it easy to work relatively to the base address of a capability.
  This instruction is not strictly necessary as it can be emulated with other instructions.
  Finally, we have the $\trestrict{}{}$ instruction which restricts the permission of a capability according to the $\le$ relation.
\end{jversion}

@@

  Additionally, when the untrusted code returns,  we need to regain access to the capabilities that represent our authority.
  To achieve this, we have to store them in a location that we can access after the return.
  However, without an encapsulation mechanism, the untrusted code would also have the same authority as us, including access to the authority we stored away.
  This is why any reasonable capability machine must have an encapsulation mechanism to allow programs to set up boundaries between security domains.

  To establish such a boundary in \trgcm{}, we can seal the return capability.
  Specifically, we can seal the return capability (which points to the instruction to be executed after the return) as a pair, together with a pointer to the stack frame where we have stashed away our capabilities (see further).
  By doing this, the untrusted code is prevented from changing the target of the return capability forcing them to return to the point we specified.
  Additionally, they do not get access to the capabilities we stashed away.
  Nevertheless, the automatic unsealing of sealed capability pairs upon invocation will ensure that the return code can still proceed as before.
  In other words, the sealing mechanism in \trgcm{} allows us to transfer control to untrusted code without giving up our capabilities or handing them over, and also control the locations in our code they can jump back to.

  It is worth pointing out that \trgcm{} does not have a direct unsealing instruction to extract a capability from its sealed version when the seal is available, but one can be (partially) emulated.
  Say you have a sealed non-executable word $\sealed{\sigma,w}$ as well as a set-of-seals capability $\seal{\sigma_\baddr,\sigma_\eaddr, \sigma}$ that contains $\sigma$, i.e.\ $\sigma \in [\sigma_\baddr,\sigma_\eaddr]$.
  It is possible to extract $w$ in $\rdata$ using $\txjmp{}{}$ in the following way.
  The idea is to use the seal to construct a sealed version of the pc capability incremented by one, and then perform an $\txjmp{}{}$ to it in combination with the sealed capability.
  This does not work for sealed executable capabilities because $\txjmp{}{}$ fails if the data capability is executable.
  However, even though we cannot extract the executable capability from $\sealed{\sigma,c}$, we can still invoke it with arbitrary arguments, since we can use the seal to construct an arbitrary data part and $\txjmp{}{}$ to the combination.\footnote{Alternatively, we could remove the current restriction in \texttt{xjmp} that requires the data capability to be non-executable but then the invoked code would often have to manually perform this check instead.}

@@

\begin{jversion}
\subsection{Decoding and encoding functions}
The operational semantics of the capability machine uses the function $\decInstr{}$ to decode instructions.
We also assume a function $\encInstr{}$ to make it easy to specify programs in terms of instructions.
Rather than defining such a decode function and an encode function, we assume that they are given with certain properties.
Most importantly, $\encInstr{} : \Instr \fun \ints$ should be the right inverse of $\decInstr{}$, i.e.
\[
\forall i \in \Instr \ldotp \decInstr{}(\encInstr{i}) = i
\]
The $\decInstr{} : \Word \fun \Instr$ should only map numbers to non-$\fail$ instructions, i.e.\ capabilities are mapped to $\fail$:
\[
  \forall c \in \Caps \ldotp \decInstr{c} = \fail
\]
These assumptions are sufficient to construct program examples in terms of instructions rather than machine words (using $\encInstr{}$), and run them on the machine (using the fact that $\decInstr{}$ is the left inverse of $\encInstr{}$).
% The $\encInstr{}$ function allows us to specify the program ``abstractly''.
% As the $\decInstr{}$ function is the left inverse of $\encInstr{}$, we can even execute the program in the operational semantics without ever worrying about what the actual encoding is.
%This is not to say that encoding is an easy problem. There are many things to get.

@@

\subsection{Components, linking, programs, and contexts}
\label{subsec:components-linking}

\begin{figure}
  \centering
  \[
    \begin{array}{rl c rl c rl}
    s &\in \Symbol & \phantom{make} &
    \mathrm{import} &\mathrel{::=} a \mapsfrom s & \phantom{make} &
    \mathrm{export} &\mathrel{::=} s \mapsto w\\
    \comp_0 & \multicolumn{7}{l}{\mathrel{::=} (\mscode,\msdata,\overline{\mathrm{import}},\overline{\mathrm{export}},\sigrets,\sigcloss,A_\linear)}\\
    \comp & \multicolumn{7}{l}{\mathrel{::=} \comp_0 \mid  (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})}
  \end{array}
\]
\vspace{0.5cm}
\begin{mathpar}
  \inference{
    \comp_0 = (\mscode[1], \msdata[1], \overline{\var{import}_1}, \overline{\var{export}_1}, \sigrets[1], \sigcloss[1],A_{\linear,1})\\
    \comp_0' = (\mscode[2], \msdata[2], \overline{\var{import}_2}, \overline{\var{export}_2}, \sigrets[2], \sigcloss[2],A_{\linear,2})\\
    \comp_0'' = (\mscode[3], \msdata[3], \overline{\var{import}_3}, \overline{\var{export}_3}, \sigrets[3], \sigcloss[3],A_{\linear,3})\\
    \mscode[3] = \mscode[1] \uplus \mscode[2] \\
    \msdata[3] = (\msdata[1] \uplus \msdata[2])[a \mapsto w \mid (a \mapsfrom s) \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}), (s \mapsto w) \in \overline{\var{export}}_3] \\
    \overline{\var{export}_3} = \overline{\var{export}_1} \cup \overline{\var{export}_2}&
    \overline{\var{import}_3} = \{ a \mapsfrom s \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}) \mid s \mapsto \_ \not\in \overline{\var{export}_3} \}\\
    \sigrets[3] = \sigrets[1] \uplus \sigrets[2] &
    \sigcloss[3] = \sigcloss[1] \uplus \sigcloss[2] &
    A_{\linear,3} = A_{\linear,1} \uplus A_{\linear,2}\\
    \dom(\mscode[3]) \mathrel{\#} \dom(\msdata[3]) & \sigrets[3] \mathrel{\#} \sigcloss[3]
  } {
    \comp_0'' = \comp_0 \bowtie \comp_0'
  }
  \and
  \inference{
    \comp_0'' = \comp_0 \bowtie \comp_0'
  }{
    \begin{aligned}
      (\comp_0'',c_{\mathrm{main},c}, c_{\mathrm{main},d}) &= \comp_0 \bowtie
      (\comp_0',c_{\mathrm{main},c}, c_{\mathrm{main},d})\\
      &=
      (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}) \bowtie \comp_0'
    \end{aligned}
  }
  \end{mathpar}
% \begin{definition}[Plugging a program into a context]
%   When $\var{comp'}$ is a context for component $\comp$ and $\comp' \bowtie \comp \rightsquigarrow \Phi$, 
%   then we write $\plug{\comp'}{\comp}$ for the execution configuration $\Phi$.
% \end{definition}
  \caption{Components and linking of components.}
  \label{fig:target-component-and-linking}
\end{figure}

The executable configuration describes the machine state, but it does not make it clear what components run on the machine and how they interact with each other.
To clarify this, we introduce notions of components and programs from which we construct executable configurations.
A component (defined in Figure~\ref{fig:target-component-and-linking}) is basically a program with entry points in the form of imports that need to be linked.
It has exports that can satisfy the imports of other components.
% Structure
A base component $\comp_0$ consists of a code memory fragment, a data memory fragment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals (see Section~\ref{sec:form-secur-with}), and a set of all the linear addresses (addresses governed by a linear capability).
The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.
An export associates a word with a symbol.
%base = library
A component is either a library component (without a main entry point) or an incomplete program with a main entry point in the form of a pair of sealed capabilities.
The latter can be seen as a program that still needs to be linked with libraries.
% Linking .
Components are combined into new components by linking them together, as long as only one has a main function.
Two components can be linked when their memories, seals, and linear addresses are disjoint.
They are combined by taking the union of each of their constituents.
For every import that is satisfied by an export of the other component, the data memory is updated to have the exported word on the imported address.
The satisfied imports are removed from the import list in the resulting linked component and the exports are obtained by combining the components' exports.

% Programs
We can now define the notion of a program as well as a context.
\begin{definition}[Programs and Contexts]
  \label{def:program-and-context}
  A \emph{program} is a component $(\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with an empty import list.
  A \emph{context} for a component $\comp$ is a component $\comp'$ such that $\comp \bowtie \comp'$ is a program.
\end{definition}
% Getting an executable from a component - postpone this to later, but mention:
% write xor exec
How a program is initialised to create an executable configuration is discussed in Section~\ref{sec:form-secur-with}.
% We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
%Further, we will only consider a specific class of well-formed components, but we will get back to this.
Some simplifications have been made in this presentation of \trgcm{}. See \citet{technical_report_popl} for details.

@@

% Informally explain how we prevent this using linear capabilities
As the examples illustrate, this naive use of memory and object capabilities does not suffice to enforce LSE and WBCF.
The problem is essentially that the stack and return pointers that a callee receives from a caller remain in effect outside their intended lifetime: either when the callee has already returned or when they have themselves invoked other code. 
Linear capabilities offer a form of revocation\footnote{Revocation in the sense that if we hand out a linear capability and later get it back, then the adversary cannot have kept a copy of it as it is non-duplicable. In other words, the adversary's access has effectively been revoked in this situation.} that can be used to prevent this from happening.

% Stack capability linear
The linear capabilities are put to use by requiring the stack capability to be linear.
On call, the caller splits the stack capability in two: one capability for their local stack frame and another one for the unused part of the stack.
The local stack frame capability is sealed and used as the data part of the sealed return pair.
The capability for the remainder of the stack is given to the callee.
% Prevention of left attack non-aliasing of linear capabilities - token like (ensuring LSE)
Because the stack capability is linear, the caller knows that the capability for their local stack frame cannot have an alias.
This means that an adversary cannot access the caller's local data because the caller has a linear capability for it and there cannot exist an alias.
The caller gives this capability to the adversary only in a sealed form, rendering it opaque and unusable.
This is illustrated in Figure~\ref{fig:stack-ptr-abuse-prev} and prevents the issue illustrated in Figure~\ref{fig:stack-ptr-abuse}.
